---
external_link: ""
date: "2019-11-01"
image:
  caption: 
links:
summary: web-based visualization
tags:
- GIS
- Data Viz
- Statistics
title: Data Viz
url_code: ""
url_pdf: ""
url_video: ""
---

```{r, warning=F, message=F, include=F}
library(pacman)
p_load(tidyverse, readr, readxl, lubridate, purrr, janitor, scales, plotly, tmap, leaflet, DT,lubridate,sf,xfun, ggridges, ggmap,tidycensus, leaflet, geojsonio, sf, tmap, tmaptools, DT, xfun,ceramic, gganimate, geojsonio,  gifski, transformr, ggspatial)
```

```{r message=FALSE, warning=FALSE, echo=F}
continuous<-read_csv("C:/Users/reidh/Documents/R/scratch/validation_traffic_counts/analysis_methodology/continuous.csv") %>%
  mutate(weekday=wday(Date, label=TRUE), month=month(Date, label=TRUE), Day1=as.character(Day))
```

We emphasis interactive data visualization because information is increasingly presented in modern, web-based forums. We utilize popular open source tools such as plotly, leaflet, and ggplot to produce visually pleasing and effective maps and charts for summarizing information.

## Distribution Plots

```{r warning=F, message=F, echo=F, out.width="150%"}
ggplotly(
continuous %>% 
  group_by(Day) %>% 
  summarise(Count=sum(Count, na.rm=T)) %>% 
  ggplot(aes(Count)) +
  geom_histogram(aes(y=..density..),colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666")  +
  theme_minimal() +
  theme(axis.title.x = element_blank(), axis.text.y=element_blank(), axis.title.y=element_blank())
)
```

## Box & Whisker Plots

```{r warning=F, message=F, echo=F, out.width="150%"}
options(scipen=999)

ggplotly(
continuous %>%
  group_by(Day,month, weekday) %>%
  summarise(Count=sum(Count,na.rm=T)) %>%
ggplot(aes(x = weekday, y = Count, color = weekday)) + 
  theme_dark() +
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank(), axis.ticks.x=element_blank()) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(labels=scales::comma)+
  facet_grid(~month) + theme(legend.position="bottom") +
  scale_color_manual(values=c("#08D6DA", "#06378D", "#BE0B05", "#F7A537", "#DEF006", "#35AA27", "#940CD8")), height=500) %>%
  layout(legend = list(orientation = "h", x = 0.0, y = -0.2)) 
```

## Raw Observations Grouped

```{r, warning=F, message=F, echo=F, out.width="130%"}
ggplotly(
  continuous %>% group_by(Day, station) %>% 
    summarise(Count=sum(Count)) %>% ggplot(aes(Day, Count, group=station, fill=station)) + 
    geom_col()+theme_minimal() + theme(axis.title.x=element_blank(), legend.position = "none", axis.title.y=element_blank()) + scale_y_continuous(labels=scales::comma) + scale_x_date() 
  )
```

## Facetted Plots

```{r warning=F, message=F, echo=F, out.width="150%"}
ggplotly(
continuous %>% 
  group_by(wday_wknd, ext_int,Day) %>% 
  summarise(Count=sum(Count, na.rm=T)) %>%
  ggplot(aes(Day,Count)) + geom_line(size=.7, color="white") +
  facet_grid(wday_wknd ~ ext_int ) +
  theme_dark() +
  theme(legend.position="bottom", legend.title = element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank(),
        axis.text.x=element_text(angle=34, hjust=1)) +
  scale_y_continuous(labels=scales::comma) )
```

## Tile Grids

```{r warning=F, message=F, echo=F, out.width="140%"}

test<-continuous %>%
  group_by(Day, weekday, month, station, ext_int) %>%
  summarise(Count=sum(Count)) %>%
  mutate(dummy="") %>%
ggplot(aes(dummy,station,  fill = Count)) + 
  geom_tile(colour = "white") + 
  facet_grid(weekday~month) + 
  scale_fill_gradient(low="red", high="green") +
  theme_dark() +
  theme(axis.ticks.x=element_blank(), axis.title.x=element_blank(), axis.title.y=element_blank(), axis.text.y=element_text(size=7), axis.text.x=element_blank(),legend.position = "none")
ggplotly(test, height = 950) %>%
  layout(legend = list(orientation = "h", x = 0.4, y = -0.2))
```

## Time Series Charts

```{r, warning=F, message=F, echo=F, out.width="130%"}
all_days<- continuous %>% 
  group_by(weekday, Day) %>% 
  summarise(Count=sum(Count, na.rm=T)) %>% 
  ungroup()

ggplotly(
  ggplot(all_days, aes(Day,Count, group=weekday, color=weekday)) + geom_line() + theme_minimal() + theme(axis.title.y = element_blank())+scale_y_continuous(labels=scales::comma)) %>%
  layout(legend = list(orientation = "h", x = 0.0, y = -0.3)) 

```

## Stacked Charts

```{r warning=F, message=F, echo=F}
ggplotly(
continuous %>% 
  group_by(Day, month, station) %>%
  summarise(Count=sum(Count,na.rm=T)) %>%
  group_by(month, station) %>%
  summarise(Count=median(Count)) %>% 
  ggplot(aes(month, Count, group=station, fill=station)) + geom_col() + theme_minimal() +
   theme(legend.title = element_blank(), axis.title.x=element_blank(),axis.title.y=element_blank(), axis.text.y=element_blank(),
        axis.text.x=element_text(size=12), legend.position = "none") 
)
```

## Interactive Maps
```{r warning=F, message=F, echo=F, out.width="130%"}
census_api_key("680398dff0a2f4c566f10c95888da7f25e55147b")
options(tigris_use_cache = TRUE)

income <- c(`Less than $10,000`= "B19001_002",
                    `$10,000 to $14,999`= "B19001_003",
                    `$15,000 to $19,999` = "B19001_004",
                    `$20,000 to $24,999` = "B19001_005",
                    `$25,000 to $29,999` = "B19001_006",
                    `$30,000 to $34,999` = "B19001_007",
                    `$35,000 to $39,999` = "B19001_008", 
                    `$40,000 to $44,999` = "B19001_009",
                    `$45,000 to $49,999` = "B19001_010",
                    `$50,000 to $59,999` = "B19001_011", 
                    `$60,000 to $74,999` = "B19001_012",
                    `$75,000 to $99,999` = "B19001_013",
                    `$100,000 to $124,999` = "B19001_014", 
                    `$125,000 to $149,999` = "B19001_015",
                    `$150,000 to $199,999` = "B19001_016", 
                    `$200,000 or more` = "B19001_017")
  dg_acs <- get_acs(geography = "tract", year=2016, 
                variables =  income, geometry=T,
                state = "NV", county="Douglas", summary_var = "B19001_001") %>%
  mutate(data_source="2016 ACS 2016 5-year Estimate", county="Douglas County")
  wa_acs <- get_acs(geography = "tract", year=2016, 
                variables =  income, geometry=T,
                state = "NV", county="Washoe", summary_var = "B19001_001") %>%
  mutate(data_source="2016 ACS 2016 5-year Estimate", county="Washoe County")
ed_acs <- get_acs(geography = "tract", year=2016, 
                variables =  income, geometry = T,
                state = "CA",county="El Dorado", summary_var = "B19001_001") %>%
  mutate(data_source="2016 ACS 2016 5-year Estimate", county="El Dorado County") 
pla_acs <- get_acs(geography = "tract", year=2016, 
                variables =  income, geometry = T,
                state = "CA",county="Placer", summary_var = "B19001_001") %>%
  mutate(data_source="2016 ACS 2016 5-year Estimate", county="Placer County") 

income1<-rownames_to_column(data.frame(income), "name") %>% dplyr::select(name) %>% pull()

all<- bind_rows( pla_acs,ed_acs,dg_acs ,wa_acs) %>%
  mutate(income2=case_when(variable=='Less than $10,000' ~10000,
                   variable== '$10,000 to $14,999'~ 12500,
                    variable=='$15,000 to $19,999' ~ 17500,
                    variable=='$20,000 to $24,999' ~ 22500,
                    variable=='$25,000 to $29,999' ~ 27500,
                    variable=='$30,000 to $34,999' ~ 32500,
                    variable=='$35,000 to $39,999' ~ 37500,
                    variable=='$40,000 to $44,999' ~ 42500,
                    variable=='$45,000 to $49,999' ~ 47500,
                    variable=='$50,000 to $59,999' ~ 55000, 
                    variable=='$60,000 to $74,999' ~ 67500,
                    variable=='$75,000 to $99,999' ~ 87500,
                    variable=='$100,000 to $124,999' ~ 112500, 
                    variable=='$125,000 to $149,999' ~ 137500,
                    variable=='$150,000 to $199,999' ~ 175000,
                    variable=='$200,000 or more'  ~ 200000),
           total=estimate*income2) %>%
  filter(total !=0) %>%
  group_by(GEOID,county) %>%
  summarise(total=sum(total), estimate=sum(estimate), average=total/estimate) %>%
   filter(!GEOID %in% c("32031940200","32031003501"))
tmap_mode("view")
tmap_leaflet(tm_shape(all) +
    tm_polygons("average", palette = "RdYlBu",title="Average Household Income") +
    tm_facets(by = "county") +
  tm_layout(legend.show=FALSE))
```

```{r}
tm_shape(all) + tm_polygons()
```

