---
date: "2020-01-11T00:00:00Z"
external_link: ""
image:
  caption: Photo by NY Times
links:
summary: Data analysis and visualization of environmental data related to the 2019 wildfires in Australia
tags:
- GIS
- Data Collection
- Data Viz
title: 2019 Australian Fires
url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""
---

During the end of 2019, wildfires proliferated across Australia. As part of the [R For Data Science TidyTuesday series](https://github.com/rfordatascience/tidytuesday), I analyzed Australian temperature and rainfall data to attempt to discern potential contributers to the historic and devastating fires. The visualizations below were created using [R](https://www.r-project.org/about.html). All code can be found on [github here](https://github.com/reidhaefer/reid.haefer/blob/master/content/post/aussie_fires/index.Rmd).


```{r warning=F, message=F, echo=F, out.width="130%", include=F}
library(pacman)
p_load(tidyverse, sf, here,mapview, lubridate, broom, ggdark,tmap, htmlwidgets, htmltools,widgetframe, leaflet)

rain <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv')

temp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv') %>%
  mutate(lon=case_when(site_name=="PERTH AIRPORT" ~ 115.9672,
                       site_name=="PORT LINCOLN AWS" ~ 135.88,
                       site_name=="KENT TOWN" ~ 138.6192,
                       site_name=="BRISBANE AERO" ~ 153.1218,
                       site_name=="SYDNEY (OBSERVATORY HILL)" ~ 151.2048,
                       site_name=="CANBERRA AIRPORT" ~ 149.1934,
                       site_name=="MELBOURNE (OLYMPIC PARK)" ~ 144.9800),
         lat=case_when(site_name=="PERTH AIRPORT" ~ 31.9385,
                       site_name=="PORT LINCOLN AWS" ~ -34.60,
                       site_name=="KENT TOWN" ~ 34.9203,
                       site_name=="BRISBANE AERO" ~ 27.3942,
                       site_name=="SYDNEY (OBSERVATORY HILL)" ~ 33.8595,
                       site_name=="CANBERRA AIRPORT" ~ 35.3052,
                       site_name=="MELBOURNE (OLYMPIC PARK)" ~ 37.8204),
         month=month(date),
         year=year(date))

fire <- sf::st_read("http://www.rfs.nsw.gov.au/feeds/majorIncidents.json")

nasa_fire <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/MODIS_C6_Australia_and_New_Zealand_7d.csv') %>%
  st_as_sf(crs=4326, coords=c("longitude","latitude"))
```

## Fire Locations

```{r warning=F, message=F, echo=F, out.width="160%"}

pal <- colorNumeric(
  palette = c('#FEBE05','#FE6005','#FE3805','#A82503','#4B1102'),
  domain = nasa_fire$brightness
)

map<-leaflet(nasa_fire) %>% 
  addCircleMarkers(fillColor = ~pal(brightness) ,color="white", weight=1, 
                   fillOpacity = 0.5, radius=6,
    popup = paste("Brightness:", nasa_fire$brightness, "<br>",
                                 "Date:", nasa_fire$acq_date, "<br>",
                                 "Time:", nasa_fire$acq_time, "<br>",
                                 "Day or Night:", nasa_fire$daynight,
                                 "NASA Confidence:", nasa_fire$confidence)) %>%
addTiles() %>% 
  addLegend("bottomright", pal = pal, values = ~brightness,
            title = "Wildfire Brightness",
            opacity = 1
  )

frameWidget(map)
```


## Are Temperatures Increasing?

A quick analysis shows that average annual temperatures have increased in almost all cities over the past 75 years.

```{r warning=F, message=F, echo=F, out.width="160%"}
temp %>%
 # filter(site_name=="PERTH AIRPORT") %>%
  group_by(year,temp_type, city_name) %>%
  summarize(temperature=max(temperature)) %>%
  ggplot(aes(year, temperature, group=temp_type, color=temp_type)) + geom_line() + geom_smooth(method = "lm") +
   facet_grid(cols = vars(city_name)) + dark_theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1),legend.position = "none",
        axis.title = element_blank()) +
  scale_colour_manual(values=c("#F7FA06","#FAA806")) +
  ggtitle("Minimum and Maximum Annual Temperatures")
```


## Less Rainfall?

The data also shows a trend for decreasing amounts of rainfall.

```{r warning=F, message=F, echo=F, out.width="130%"}
rain %>%
  group_by(year) %>%
  summarize(rainfall=mean(rainfall,na.rm=T)) %>%
  ggplot(aes(year, rainfall)) + geom_line(colour="#06EDFA") + geom_smooth(method = "lm", colour="white") +
dark_theme_bw() + theme(axis.text.x=element_text(angle=45, hjust=1),axis.title = element_blank()) +
  ggtitle("Average Annual Rainfall")
```