---
date: "2019-10-29T00:00:00Z"
external_link: ""
image:
  caption: Photo by Reid Haefer
links:
summary: Analysis of Central Park Squirrel Census dataset as part of the October 29th, 2019 TidyTuesday data series.
tags:
- GIS
- Data Collection
- Data Viz
title: NYC Squirrels
url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""
---

As part of the [R For Data Science TidyTuesday series](https://github.com/rfordatascience/tidytuesday), I analyzed data from the first annual [Central Park Squirrel Census](https://www.thesquirrelcensus.com/). The 2018 [survey](https://www.citylab.com/life/2019/06/squirrel-census-results-population-central-park-nyc/592162/) identified over 2,000 individual squirrels in New York City's Central Park and documented detailed individual characteristics such as location, age, color & behavior. 


```{r warning=F, message=F, echo=F, out.width="130%"}
library(pacman)
p_load(sf, tidyverse,mapview, tmap,ggmap, ggspatial, mapdata, viridis)

register_google(key = "AIzaSyBBVBD3Y_tifwVFsC0sWq6vlK44RXlM-i8")

## What causes squirrels to act the way they do?
# AM vs PM and day of week
# Location
# Age
# fur color
# on ground/above ground/ distance above ground

nyc <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-29/nyc_squirrels.csv") %>%
  mutate(date=as.Date(as.character(date), "%m%d%Y"),
         above_ground_sighter_measurement=as.numeric(na_if(above_ground_sighter_measurement,FALSE)),
         tree_ground=case_when(location== "Above Ground" ~ "Tree-Dwelling Squirrel",
                               location== "Ground Plane" ~ "Ground-Dwelling Squirrel"))
```


## Where are the most squirrels in Central Park?

```{r warning=F, message=F, echo=F, out.width="130%"}
nyc_s <- nyc %>% filter(!is.na(location))
p<-ggmap(get_map(location = c(lon = -73.968285, lat =40.785091 ),
                    zoom = 14, scale = 2,"satellite"))

p +
  coord_cartesian() +
  geom_hex( data = nyc_s, aes(x=long, y=lat),bins = 40, alpha=.7, color="white") +
  scale_fill_viridis(option="A") +
  theme_bw() +
  theme(axis.text=element_blank(), axis.title = element_blank(), axis.ticks = element_blank()) +
  ggtitle("2018 Central Park Squirrel Census - Count Density") +
  labs(fill = '# of Squirrels')
```


## What are other characteristics of Central Park squirrels?

```{r warning=F, message=F, echo=F, out.width="130%"}
nyc_s <- nyc %>% filter(!is.na(location)) %>% filter(!is.na(primary_fur_color))
g<-ggmap(get_map(location = c(lon = -73.968285, lat =40.785091 ),
                 zoom = 14, scale = 2,"toner-lite"))

g +
  geom_point(data=nyc_s,aes(x=long,y=lat,fill=primary_fur_color), show.legend = T,colour="white",pch=21, size=2) +
  facet_grid(shift ~ tree_ground) +
  scale_fill_manual(values=c("black","#8D4F0D","#A6A5A4")) +
  theme_dark() +
  theme(axis.text=element_blank(),axis.title=element_blank() ,axis.ticks = element_blank())  +
  labs(fill = 'Squirrel Color')
```