---
date: "2020-01-25T00:00:00Z"
external_link: ""
image:
  caption: Photo from www.kaylinpavlik.com
links:
summary: Analysis of music music genres and characteristics over time using a large Spotify database.
tags:
- Data Science
- Data Viz
- Big Data
title: Music Characterization
url_code: ""
url_pdf: ""
url_slides: ""
url_video: ""
---

Has popular music evolved over time?. As part of the [R For Data Science TidyTuesday series](https://github.com/rfordatascience/tidytuesday), I analyzed a large dataset of Spotify songs, which included detailed characteristics of each song. The visualizations below were created using [R](https://www.r-project.org/about.html). All code can be found on [github here](https://github.com/reidhaefer/reid.haefer/blob/master/content/post/spotify/index.Rmd).


```{r warning=F, message=F, echo=F, out.width="130%", include=F}
library(pacman)
p_load(tidyverse, broom, lubridate, plotly, gganimate, formattable)

spot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv') %>%
  mutate(release_date=as.Date(track_album_release_date, "%Y-%m-%d"),
         year=year(release_date)) %>%
  pivot_longer(cols=c(12:23), names_to="character", values_to = "values" )

spot_wide <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv') %>%
  mutate(release_date=as.Date(track_album_release_date, "%Y-%m-%d"),
         year=year(release_date)) 
```

## All genres and characteristics over time

```{r warning=F, message=F, echo=F, fig.width=8, fig.height=10}
a<-spot %>%
  filter(year >1980 & !character %in% c("valence", "mode","key","liveness")) %>%
  group_by(year, character, playlist_genre) %>%
  summarise(values=mean(values,na.rm=T)) %>%
  na.omit() %>%
  ggplot(aes(year, values )) + geom_line(color= "#069A89", size=1) + 
  #geom_smooth(aes(color="#F0A13B")) +
  facet_grid(rows= vars(character), scales="free", cols = vars(playlist_genre)) +
  theme_classic() +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.title = element_blank(), legend.position = "none",
        axis.text.x=element_text(angle=45, hjust=1, size=7)) 
a +
  transition_reveal(year) + shadow_mark()

spot %>%
  filter(year >1980 & !character %in% c("valence", "mode","key","liveness")) %>%
  group_by(year, character, playlist_genre) %>%
  summarise(values=mean(values,na.rm=T)) %>%
  ggplot(aes(year, values )) + geom_line() + geom_smooth(aes(color="#F0A13B")) +
  facet_grid(rows= vars(character), scales="free", cols = vars(playlist_genre)) +
  theme_classic() +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.title = element_blank(), legend.position = "none",
        axis.text.x=element_text(angle=45, hjust=1, size=7)) 
```

```{r warning=F, message=F, echo=F, fig.width=9, fig.height=5}
b<-spot %>%
  filter(year >1980 & !character %in% c("valence", "mode","key","liveness")) %>%
  filter(playlist_genre=="rap" & character=="danceability") %>%
  group_by(year, character, playlist_genre) %>%
  summarise(values=mean(values,na.rm=T)) %>%
  na.omit() %>%
  ggplot(aes(year, values)) + geom_line( size=2, color="#069A89") + 
  #geom_smooth(aes(color="#F0A13B")) +
  theme_minimal() +
  ggtitle("Rap Song Danceability Over Time") +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.title = element_blank(), legend.position = "none", axis.text.x=element_text(angle=45, hjust=1, size=14)) 
b +
  transition_reveal(year) + shadow_mark()
```

```{r warning=F, message=F, echo=F, fig.width=9, fig.height=5}
c<-spot %>%
  filter(year >1980 & !character %in% c("valence", "mode","key","liveness")) %>%
  filter(playlist_genre=="rock" & character=="loudness") %>%
  group_by(year, character, playlist_genre) %>%
  summarise(values=mean(values,na.rm=T)) %>%
  na.omit() %>%
  ggplot(aes(year, values)) + geom_line( size=2, color="#069A89") + 
  #geom_smooth(aes(color="#F0A13B")) +
  theme_minimal() +
  ggtitle("Rock Song Loudness Over Time") +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(), axis.title = element_blank(), legend.position = "none", axis.text.x=element_text(angle=45, hjust=1, size=14)) 
c +
  transition_reveal(year) + shadow_mark()
```

## Linear Model

What determines a song's popularity? A generalized linear model can help us understand.

track popularity ~ danceability + energy + key + loudness + mode + speechiness + acousticness + instrumentalness + liveness + valence + tempo + duration

```{r warning=F, message=F, echo=F, out.width="160%"}
glm_fit <- glm(track_popularity ~ 
            danceability + 
              energy + 
              key + 
              loudness + 
              mode + 
              speechiness + 
              acousticness + 
              instrumentalness + 
              liveness + 
              valence + 
              tempo + 
              duration_ms, 
          data=spot_wide) 

formattable(tidy(glm_fit) %>%
              mutate(estimate=round(estimate,3),
                     std.error=round(std.error,3),
                     statistic=round(statistic,3),
                     p.value=round(p.value,4)), 
            #align = c("l",rep("r", NCOL(prevalence) - 1)),
            list(`term` = formatter("span", style = ~ style(color = "grey", font.weight = "bold")), 
                 `estimate` = formatter("span", style = x ~ style(color = ifelse(x < 0, "red", "#3B7CF0"), font.weight = "bold")),
                 `std.error` = color_bar("#F0C93B"), 
                 `statistic` = formatter("span", style = x ~ style(color = ifelse(x < 0, "red", "#3B7CF0"), font.weight = "bold"))))


```

